---
title: "GenBank record parsing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{parsing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Introduction

In this vigennette we will demonstrate how the `gb_extract()` function works. Restez downloads and stores all GenBank records in text format. Ordinarily, to be able to extract relevant bits of information from a text record in a systematic way we would need to make use of [regular expressions](https://en.wikipedia.org/wiki/Regular_expression). `gb_extract()` uses regular expressions, so we don't have to. Here's how it works.

#A GenBank record

restez comes with an example GenBank record, AY952423, which can be viewed [online](https://www.ncbi.nlm.nih.gov/nuccore/AY952423.1). We can retrieve the record with `rentrez` or load it using the `data()`. We can visual the record as it appears online using `cat()` -- like print but parses the newline spaces correctly.

```{r record, echo=TRUE}
library(restez)
# record <- rentrez::entrez_fetch(db = 'nucleotide', id = 'AY952423', rettype = 'gb', retmode = 'text')
data(record)
cat(record)
```

#Extracting

We can extract different elements of the above record with `gb_extract()` ....

```{r extract, echo=TRUE}
# such as the LOCUS information ...
(gb_extract(record = record, what = 'locus'))
# the accession
(gb_extract(record = record, what = 'accession'))
# the accession + version
(gb_extract(record = record, what = 'version'))
# the organism name
(gb_extract(record = record, what = 'organism'))
# the sequence definition line
(gb_extract(record = record, what = 'definition'))
# the keywords (this record doesn't have any ....)
(gb_extract(record = record, what = 'keywords'))
# even the features as a list object
features <- gb_extract(record = record, what = 'features')
print(features[[1]])
# and of coure the sequence itself
seq <- gb_extract(record = record, what = 'sequence')
str(seq)
```

#From the database

You can try and the above functions out manually by downloading them through the [`rentrez` package](https://github.com/ropensci/rentrez). You can also parse the text formatted records from a restez database.

```{r database, eval=FALSE, include=TRUE}
# ensure restez database is already set-up
record <- gb_record_get(id = 'ACCESSIONID')
gb_extract(record = record, what = 'features')
```
