---
title: "Create and Query a Local Copy of GenBank in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{restez_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# hide line spinner
assignInNamespace(x = 'custom_download2', value = restez:::custom_download,
                  ns = 'restez')
```

#1. Download GenBank

In the first section of this introduction to `restez` we will explain how to download parts of GenBank on to your computer and from these downloads, construct a local database. GenBank can be downloaded from NCBI using the File Transfer Protocol (FTP), the online portal for these downloads can be viewed from your web browser,
[FTP NCBI GenBank](ftp://ftp.ncbi.nlm.nih.gov/genbank/). GenBank is hosted at this portal as 1000s of compressed sequence files called flatfiles. New versions of these flatfiles are released once a month as new sequence information is uploaded. For information on the latest GenBank release, see the [NCBI statistics page](https://www.ncbi.nlm.nih.gov/genbank/statistics/).

`restez` downloads a selection of these flatfiles and unpacks them into an SQL-like database (specifically, [MonetDB](https://www.monetdb.org/)) that can be queried using `restez`'s functions. Because there are potentially huge amounts of sequence information, all of which is unlikely to be of interest to a user, `restez` allows you to select the sets of flatfiles to download based on GenBank's domains. For the most part these sets of flatfile types are determined by taxonomy, for example, a user can download all primate or plant sequences.

Here, we will explain how to set up `restez`, download genbank and then create a database from the downloaded files.

##1.1 Setting up

Before we can download anything, we need to tell `restez` where we would like to store the downloaded files and the eventual GenBank database. To do this we use the `restez_path_set` function. With this function a user can provide a system file path to a folder within which the function will create a new folder called 'restez'. In this example, we will create a `restez` folder in a temporary folder.

```{r presetup, include=FALSE}
rstz_pth <- tempdir()
restez::restez_path_set(filepath = rstz_pth)
restez::db_delete(everything = TRUE)
```
```{r setting up, echo=TRUE}
library(restez)
# rstz_pth <- tempdir()
restez_path_set(filepath = rstz_pth)
```

A user must always set the `restez` path in order to use the `restez` library and we are reminded to do this whenever we load the package. (We are also reminded to run `restez_connect` but we'll come to that a bit later.) Additionally, by running `restez_path_set` on a new folder messages are printed to console telling us that the path was created and that a downloads subfolder was created.

> **Note**, whenever you intend to use `restez` you will need to specify the `restez` path. If you are using it regularly, it might be a good idea to set it using your .Rprofile.

##1.2 Download

Now we can download GenBank files using `db_download()`. This is an interactive function, it looks up the latest GenBank release, parses the release notes and prints to console the available sets of flatfiles that can be downloaded. For this example, we will select the smallest avaiable domain which is 'Unannotated' and can be pre-specified with '20'. Normally, you would run this function without the `preselection` argument and just wait for the function to prompt you for a domain selection.

After launching, the download function will ask you whether it is OK to download and set up a database for the selected sequence files. If the selection is likely to be large you can always quit the process using either `Esc` or `Ctrl+c`. Please note that the diskspace estimates are preliminary. The package trys to be conservative in its estimates.

```{r download, echo=TRUE}
db_download(preselection = '20')
```

Now the download has begun, we just need to wait for it to finish. For the largest domains, this can take quite a while and may be prone to server errors. If you ever encounter a failed downloaded file it is possible to re-run the `db_download` function again and -- so long as the same selection is made -- the function will only download the missing files.

##1.3 Create a database

After the download has completed, we need to use `db_create()` to create our local database. This looks in the downloads folder of our `restez` path, breaks these files up into separate GenBank records and adds them to the SQL-like database. Again, for very large collections of sequences this can take quite a while.

Before running the `db_create` function, however, we must first establish a connection to the database in the `restez` path using the `restez_connect` function. As with SQL databases we must always make sure to connect and then disconnect after we have interacted with a database.

```{r create database, echo=TRUE}
restez_connect()
db_create()
restez_disconnect()
```

`db_create()` allows a user to specify minimum and maximum sequence lengths. It's always a good idea to limit the number of sequences in a database so that look-up times are faster. If you know you are only intersted in certain lengths of sequences it is a good idea to limit the sequences in the database at this stage. You can always run `db_create()` again to change the limits. You will simply need to delete the oirignal database first with `db_delete`.

##1.4 Checking the setup

After the download and the database creation steps are complete, we can confirm our setup using `restez_status`.

```{r confirm setup, echo=TRUE}
library(restez)
restez_path_set(rstz_pth)
restez_connect()
restez_status()
restez_disconnect()
```

The status function allows a user to always touch base with their `restez` set-up. It can be run at any point, before and/or after download or database creation. It is also designed to provide useful messages to a user for what they need to do next in order for them to make queries to their database. If you ever get confused, run `restez_status` to see what is happening.

Additionally, if you are developing your own functions that depend on `restez`, you can use `restez_ready` which will simply return TRUE or FALSE on whether the database can be queried.

```{r confirm setup2, echo=TRUE}
library(restez)
restez_path_set(rstz_pth)
restez_connect()
if (restez_ready()) {
  print('Database is ready to be queried!')
} else {
  print('Database cannot be queried :-(')
}
restez_disconnect()
```

#2. Query

Once a restez database has been set up we can query the database using the `gb_*_get()` functions. These functions allow us to retreive specific coulmns in the SQL-like database: 'sequence', 'definition', 'accession', 'version' and 'organism'. Also, they allow us to get the whole text formatted record and sequence data in fasta format. In this example, we can use `list_db_ids()` to identify accession numbers in the database. We could also use `entrez_search()` provided the database contains sequences of interest to us, see ['search and fetch'](https://antonellilab.github.io/restez/articles/search_and_fetch.html).

> Always remember to disconnect after you have finished your querying!

```{r query, echo=TRUE}
library(restez)
restez_path_set(rstz_pth)
restez_connect()
ids <- suppressWarnings(list_db_ids(db = 'nucleotide', n = 100))
(id <- sample(ids, 1))
# sequence
str(gb_sequence_get(id))
# definition
(gb_definition_get(id))
# version
(gb_version_get(id))
# organism
(gb_organism_get(id))
# fasta
cat(gb_fasta_get(id))
# Note, for large databases these requests can take a long time to complete.
# Always try and limit the size of the database by specifying min and max
# sequence lengths with db_create()
# Also note, if an id is not present in the database nothing is returned
cat(gb_fasta_get(id = c(id, 'notanid')))
# Always rememeber to disconnect
restez_disconnect()
```

Additionally, for more flexibility and options for extracting sequence record information see [GenBank record parsing](https://antonellilab.github.io/restez/articles/parsing.html).

#3. Entrez

Entrez wrappers are part of the `restez` package. These allow a user to make use of the local GenBank using functions that were built for [`rentrez`](https://github.com/ropensci/rentrez). This minimises the amount of coding changes required for any Entrez dependent code.

>Currently, only `entrez_fetch()` is available with restez and only text formatted rettypes are allowed.

```{r entrez, echo=TRUE}
library(restez)
restez_path_set(rstz_pth)
restez_connect()
ids <- suppressWarnings(list_db_ids(db = 'nucleotide', n = 100))
(id <- sample(ids, 1))
# get fasta sequences with entrez
res <- restez::entrez_fetch(db = 'nucleotide', id = id, rettype = 'fasta')
cat(res)
# entrez_fetch will also search via Entrez for any ids not in the db
plant_sequence <- 'AY952423'
res <- restez::entrez_fetch(db = 'nucleotide', id = c(id, plant_sequence),
                            rettype = 'fasta')
cat(res)
restez_disconnect()
```

#4. More information

For more information about `restez` see the other tutorials:

1. [Build a database of all rodents](https://antonellilab.github.io/restez/articles/1_rodents.html)
2. [How to search for and fetch sequences](https://antonellilab.github.io/restez/articles/2_search_and_fetch.html)
3. [Advanced parsing of a GenBank record](https://antonellilab.github.io/restez/articles/3_parsing.html)
4. [Running phylotaR with restez](https://antonellilab.github.io/restez/articles/4_phylotar.html)
5. [Tips and tricks](https://antonellilab.github.io/restez/articles/5_tips_and_tricks.html)
